---
title: POL 201 - Methods and Statistics
subtitle: Descriptive Statistics
author: Andrew O'Geen
institute: Department of Political Science, Davidson College
output: 
  beamer_presentation:
    keep_tex: true
    latex_engine: xelatex
    template: rmd_teaching_slides.tex

---

## Descriptive v. Inferential

$$
\begin{aligned}
\text{Descriptive Statistics} &= 
\begin{cases}
    \text{Frequencies} \\
    \text{Basic measurments}\\
    \text{How many? How much?}
\end{cases}
\\
\\
\text{Inferential Statistics} &= 
\begin{cases}
    \text{Hypothesis testing} \\
    \text{Confidence intervals}\\
    \text{Associations between phenomena}\\
    \text{Generalizable conclusions}
\end{cases}
\end{aligned}
$$ 

## {.c}

\centering
Why do we need descriptive statistics?


## Dispersion

```{r, echo=F}
x   <- seq(5,15,length=1000)
y   <- dnorm(x,mean=10, sd=1)
z   <- dnorm(x,mean=10, sd=5)
par(mfrow=c(1,2))
plot(x,y, type="l", lwd=1, main="Standard Deviation=1")
plot(x,z, type="l", lwd=1, main="Standard Deviation=5")
``` 
 

## The Mode

 - Most basic measure of central tendency
 - The value that we see most often in the data
 

## The Mode

 - Most basic measure of central tendency
 - The value that we see most often in the data
 - What is the mode of the following set of numbers?  
 `[14, 10, 10, 39, 21, 4, 6, 94, 10]` 
 
## The Mode

```{r comment=NA}
# Create variable containing first set of numbers
z <- c(14, 10, 10, 39, 21, 4, 6, 94, 10)
# Generate table to see the modal value
table(z)
```

## The Mode

 - Most basic measure of central tendency
 - The value that we see most often in the data
 - What is the mode of the following set of numbers?  
 `[14, 10, 10, 39, 21, 4, 6, 94, 10]`  
 - What about this set? `[1, 2, 2, 2, 3, 4, 5, 5, 5]`

## The Mode

```{r comment=NA}
# Create variable containing first set of numbers
b <- c(1, 2, 2, 2, 3, 4, 5, 5, 5)
# Generate table to see the modal value
table(b)
```

## The Mean

$$\large mean_x =  \frac{x_{1} + x_{2} + x_{3} + . . . + x_{n} }{n}$$
Or, more formally
 
$$\large \bar{x} = \frac{\sum\limits_{i=1}^n x_i}{n}$$ 

## The Mean

 - Different notation for populations and samples
 - True for all statistics
 - mean of the population: $\mu_x$ 
 - mean of sample: $\bar{x}$ 
 
## The Mean

```{r echo=FALSE}
set.seed(123456)
```

```{r comment=NA}
# Generate random variable with 1000 observations
x <- rnorm(1000, mean=24, sd=2)
# Calculate the mean
mean(x)
```

## The Median

 - Mean can be misleading
 - Bill Gates walks into a bar...
 - Middle of the data. Equal number of observations above and below.
 
## The Median {.c}

$$
\Large  \tilde{x} = 
\begin{cases}
    x_{(\frac{n+1}{2})}              & \text{when n is odd} \\
    \frac{1}{2}({x_{(\frac{n}{2})}+x_{(1+\frac{n}{2})})}  & \text{when n is even}
\end{cases}
$$ 

## The Median

```{r comment=NA}
median(x)
x[1001] <- 48000
mean(x)
median(x)
```

## The Interquartile Range

 - Bounds defined by the 25th and 75th percentiles
 - 25% of the data are less than or equal
 - 75% of the data are less than or equal  

$$\LARGE IQR_x = [x_{(\frac{n}{4})}, x_{(\frac{3n}{4})}]$$

## The Interquartile Range

 
```{r comment=NA}
summary(x)
IQR(x)
25.33-22.70
```

## The End {.c}
\centering
What questions do you have?


