---
title: POL 201 - Methods and Statistics
subtitle: Descriptive Statistics, part II
author: Andrew O'Geen
institute: Department of Political Science, Davidson College
output: 
  beamer_presentation:
    keep_tex: true
    latex_engine: xelatex
    template: rmd_teaching_slides.tex

---

```{r eval=T,echo=F,message=F}
load("data/pres12.rda")
```

## The Basics

 - An important part of statistical modeling is understanding and explaining variation.

 - Start simple.  Describe variation in a single variable as simply as possible.

 - Measures of central tendency don't really speak to variation.


## The Basics

- Think of central tendency as very simple model of data

- Each observation on a variable then has two components, the model value and the variation from the model

- Individual case = model value + deviation (aka "residual") 

## Variance 


Height    |   Model Value |    Residual    |  Squared Residual
----------+---------------+----------------+-----------------
72.0      |   66.1        |   5.9          |   34.81   
62.0      |   66.1        |   -4.1         |   16.81  
68.0      |   66.1        |   1.9          |   3.61   
65.5      |   66.1        |   -0.6         |   0.36   
63.0      |   66.1        |   -3.1         |   9.61  

Sum of Squares =  $\Sigma(x-\bar{x})^2$ = 65.2

## Variance and Standard Deviation 

- To calculate the variance, we simply divide the sum of squares by $n-1$

$$\LARGE var_{x} = \frac{\Sigma(x-\bar{x})^2}{n-1}$$

- But the units are funky.  Here it is squared inches.

- We need something more useful...

## Standard Deviation

- Calculating the standard deviation is easy!
- $\sigma$ versus $S_{X}$

$$
\LARGE S_{x} = \sqrt{\frac{\Sigma(x-\bar{x})^2}{n-1}}
$$

- By taking the square root, we make the units the standard units in the data

- The standard deviation gives us a rough estimate of the typical deviation from the mean
 
## What was Obama's vote share by state in 2012?

```{r comment=NA}
mean(pres12$Obama12Share)
var(pres12$Obama12Share)
sd(pres12$Obama12Share)
sqrt(var(pres12$Obama12Share))
```

## In Pictures

```{r, echo=F}
x   <- seq(5,15,length=1000)
y   <- dnorm(x,mean=10, sd=1)
z   <- dnorm(x,mean=10, sd=5)
par(mfrow=c(1,2))
plot(x,y, type="l", lwd=1, main="Standard Deviation = 1")
plot(x,z, type="l", lwd=1, main="Standard Deviation = 5")
```

## For later... 

 - Explaining variation
 - Dividing variance: model v. residual
 - What percent of the total does our model capture?
 - How well are we doing?

## The End {.c}
\centering
What questions do you have?


